### 2022-06-23 presentation/presentation.tex
working on the presentation

\documentclass{beamer}

% Style:
\usepackage{fontspec}
\setsansfont{Candara}
\usetheme{Rochester}
\definecolor{primarycolor}{RGB}{15,20,40}
\usecolortheme[named=primarycolor]{structure}
\setbeamercolor{normal text}{fg=primarycolor}

\title{\textsc{\bfseries Transfer Learning with\\Vision Transformers}}
\author{\textbf{Vincent} Tonkes}
\institute{\textbf{University} of Groningen}
\date{}

% Also makes it bold and small caps
\newcommand{\createtitle}[1]{\frametitle{\textsc{\bfseries#1}}}

\begin{document}

{
\setbeamercolor*{frametitle}{bg=white}
\frame{
\titlepage
{\tiny Supervised by:}\\\small\textbf{Matthia} Sabatelli\tiny, PhD
}
}

% Leaves them out on slide two and highlights them on slide 3
\newcommand{\toexplain}[1]{
    \textcolor<2>{white}{\textcolor<3>{red}{#1}}
}

\begin{frame}
\createtitle{Research Question}
\begin{center}
\textbf{\large ``\textit{To what extent (and in what manners) does utilizing \toexplain{Transfer Learning} allow \toexplain{Vision Transformers} to perform on par with \toexplain{Convolutional Neural Networks} on common \toexplain{art classification problems}, when compute resources are limited to a single \toexplain{GPU}?}''}
\end{center}
\end{frame}

\end{document}