### 2022-06-23 writing/report.tex

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage{bachelorproject}

\title{\vspace{-0.5in}{\bfseries\scshape 
    Transfer Learning for computationally limited training of Vision Transformers on art classification tasks
    } \\ 
    \vspace{1ex} 
    \normalsize{Bachelor's Project Thesis}\vspace{-2ex}
}

\author{
    \normalsize{V. Tonkes, s3617157, v.tonkes@student.rug.nl,} \\ 
    \normalsize{Supervisor: M. Sabatelli, PhD}
}

\date{\vspace{-5ex}}

\begin{document}

\twocolumn[
  \maketitle
  \begin{@twocolumnfalse}
%   \input{./tex/abstract.tex}
  \end{@twocolumnfalse}
]

\thispagestyle{firststyle}

\input{tex/draft_introduction.tex}
\input{tex/draft_methods.tex}
\input{tex/draft_results.tex}

\bibliographystyle{apacite}
\bibliography{literature}

\clearpage

%\begin{appendices}
%    \input{tex/appendixA.tex}
%    \newpage
%    \input{tex/appendixB.tex}
%    \newpage
%\end{appendices}

\end{document}