### 2022-06-26 presentation/presentation.tex
Slideshow done. First draft at least

\documentclass{beamer}

% Style:
\usepackage{fontspec}
\setsansfont{Candara}
\usetheme{Rochester}
\definecolor{primarycolor}{RGB}{15,20,40}
\usecolortheme[named=primarycolor]{structure}
\setbeamercolor{normal text}{fg=primarycolor}
\setbeamertemplate{caption}{\raggedright\insertcaption\par}
\setbeamertemplate{bibliography item}[text]

% For tables:
\usepackage{colortbl}
\definecolor{bestcol}{RGB}{99, 255, 105}
\definecolor{secondbestcol}{RGB}{250, 255, 99}
\definecolor{worstcol}{RGB}{255, 141, 99}
\usepackage{arydshln}

\title{\textsc{\bfseries Transfer Learning with\\Vision Transformers}}
\author{\textbf{Vincent} Tonkes}
\institute{\textbf{University} of Groningen}
\date{}

% Also makes it bold and small caps
\newcommand{\createtitle}[1]{\frametitle{\textsc{\bfseries#1}}}

% Make \cite commands tiny
\newcommand{\tinycite}[1]{{\tiny \cite{#1}}}

\begin{document}

%-------------------------------------------------------------------------------
% Leaves them out on slide two and highlights them on slide 3
{
\setbeamercolor*{frametitle}{bg=white}
\frame{
\titlepage
{\tiny Supervised by:}\\\small\textbf{Matthia} Sabatelli\tiny, PhD
}
}

%-------------------------------------------------------------------------------

\input{tex/question.tex}
\input{tex/answer.tex}

%-------------------------------------------------------------------------------
{
\setbeamercolor*{frametitle}{bg=white}
\begin{frame}
\centering
\textsc{\Huge \textbf{That's} all!}

\pause
{\small except for the questions :-)}
\end{frame}
}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}[t, allowframebreaks]
\createtitle{References}
\bibliographystyle{acm}
{\tiny
\bibliography{literature}}
\end{frame}
%-------------------------------------------------------------------------------

\end{document}