### 2022-07-13 writing/report.tex
Slowly getting there.. Still working on the 'discussion' section. Also started writing the conclusion.

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage{bachelorproject}

\title{\vspace{-0.5in}{\bfseries\scshape 
    Transfer Learning for computationally limited training of Vision Transformers on art classification tasks
    } \\ 
    \vspace{1ex} 
    \normalsize{Bachelor's Project Thesis}\vspace{-2ex}
}

\author{
    \normalsize{V. Tonkes, s3617157, v.tonkes@student.rug.nl,} \\ 
    \normalsize{Supervisor: M. Sabatelli, PhD}
}

\date{\vspace{-5ex}}

\begin{document}

\twocolumn[
  \maketitle
  \begin{@twocolumnfalse}
%   \input{./tex/abstract.tex}
  \end{@twocolumnfalse}
]

\thispagestyle{firststyle}

\input{tex/draft_introduction.tex}
\input{tex/draft_methods.tex}
%\input{tex/draft_results.tex}
\input{tex/draft_experiments.tex}
\input{tex/draft_conclusion.tex}

\section*{Acknowledgements}
\textit{The author would like to thank {the Center for Information Technology of the University of Groningen} for their support and for providing access to the Peregrine high performance computing cluster. He would also like to thank {dr. M. Sabatelli} for his guidance during this project.}

%\clearpage

\bibliographystyle{apacite}
\bibliography{literature}

%\clearpage

%\begin{appendices}
%    \input{tex/appendixA.tex}
%    \newpage
%    \input{tex/appendixB.tex}
%    \newpage
%\end{appendices}

\end{document}